---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# classiFunc


## Overview
The classiFunc package implements methods for functional data classification. The main functions of this package are classiKnn, a k nearest neighbor estimator for functional data, and classiKernel, a kernel estimator for functional data. The package uses efficiently implemented semimetrics to create the distance matrix of the functional observations in the function computeDistMat. 

## Using classiFunc
For installation instructions, see below.
A hands on introduction to \code{classiFunc} can be found in the [vignette](https://cran.r-project.org/web/packages/classiFunc/vignettes/classiFunc.html).
Details on specific functions are in the [reference manual](https://cran.r-project.org/web/packages/classiFunc/classiFunc.pdf).

## Issues & Feature Requests
For issues, bugs, feature requests etc. please use the [Github Issues](https://github.com/maierhofert/classiFunc/issues). Input is always welcome.

## Installation

You can install the current classiFunc version from CRAN with:

```{r cran-installation, eval = FALSE}
install.packages("classiFunc")
```


or the latest patched version from Github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("maierhofert/classiFunc")
```

<!-- ## Example -->

<!-- This is a basic example which shows you how to solve a common problem: -->

<!-- ```{r example} -->
<!-- ## basic example code -->
<!-- ``` -->
